Mergeable Types

### 2017-02-28 paper/meta.tex
Minor edits in meta.tex

Fortunately, unlike source control systems where branching structure
is entirely dictated by the user, \name abstracts away branching
structure from the programmer, and hence retains the ability to
manifest it in a way that it deems fit. In particular, \name solves
the problem of multiple LCAs by suitably constraining the branching
structure such that the problem never arises. The constraints are
imposed either implicitly, as a result of how the operational semantics
defines an atomic step, or explicitly, by insisting that certain
conditions be met before merging a pair of versions
(\rulelabel{E-Pull-Wait}). First, the operational semantics already
disables criss-cross merges since it only ever merges versions that
are latest on their respective branches. Second, we impose certain
pre-conditions on the merging branches to preempt the structure shown
in Fig.~\ref{fig:external-lcas}. The intuition is as follows: consider
the branch $b_3$ at the instance of merging $v_{12}$. Since it has
already merged $v_{22}$, $v_{22}$ could be a common ancestor for $b_3$
and some other branch (call it $b$). Now, if $b_3$ merges $v_{12}$,
the same could be true of $v_{12}$. Since $v_{12}$ and $v_{22}$ are not
orderd by the ancestor relation, both become LCAs of
$b_3$ and $b$. We observe that this scenario can be prevented if, when
merging $b_1$, $b_3$ insists on an ancestor relation between the last
merged version ($v_{22}$) and the currently merging version.  We call
the last merged version of a branch its \emph{locus}. By maintaining a
total order of among locii of a branch $b$ as new versions are
created, we effectively guarantee the presence of a version (the
locus) that is lower than all the ancestors of the latest version on
$b$. Next, by requiring the locus to be an ancestor of the merging
version, we ensure that all the common ancestors have a single lowest
element, thus enforcing the uniqueness of the LCA.

@@

If $v_1$ and $v_2$ referred in the above definition are the latest
versions on their respective branches $b_1$ and $b_2$, we say $b_2$ is
mergeable into $b_1$, or more generally, $b_1$ and $b_2$ are
mergeable. The definition essentially requires all of $v_1$'s causal
history that preceeds its locus $v_o$ to be included in $v_2$'s causal
history, and none of $v_1$'s history that succeeds $v_o$ to be
included in $v_2$'s history. This allows us to view versions $v_1$ and
$v_2$ as having been independently evolved from a common causal
history whose maximum (w.r.t the ancestor relation) is the version
$v_o$. Consquently, $v_o$ becomes the LCA for $v_2$'s merge into
$v_1$. Generalizing this observation for the latest versions on
any two branches, we state the following theorem:

\begin{theorem} [\bfseries Unique LCA]
Every pair of branches in a legal history $H$, if mergeable as per
Def.~\ref{def:mergeability}, have a unique least common ancestor.
\end{theorem}

@@

\begin{theorem} [\bfseries Progress]
In a legal branching history $H$ produced by the operational
semantics, if two branches, $b_i$ and $b_j$ are not mergeable (as per
Def.~\ref{def:mergeability}), then there exists a sequence of fork and
merge operations (between mergeable versions) that can be performed on
$H$ to yeild a new history $H'$, where $b_i$ and $b_j$ are mergeable.
\end{theorem}

>>>>

Fortunately, unlike source control systems where branching structure
is entirely dictated by the user, \name abstracts away branching
structure from the programmer, and hence retains the ability to
manifest it in a way that it deems fit. In particular, \name solves
the problem of multiple LCAs by suitably constraining the branching
structure such that the problem never arises. The constraints are
imposed either implicitly, as a result of how the operational semantics
defines an atomic step, or explicitly, by insisting that certain
conditions be met before merging a pair of versions
(\rulelabel{E-Pull-Wait}). First, the operational semantics already
disables criss-cross merges since it only ever merges versions that
are latest on their respective branches. Second, we impose certain
pre-conditions on the merging branches to preempt the structure shown
in Fig.~\ref{fig:external-lcas}. The intuition is as follows: consider
the branch $b_3$ at the instance of merging $v_{12}$. Since it has
already merged $v_{22}$, $v_{22}$ could be a common ancestor for $b_3$
and some other branch (call it $b$). Now, if $b_3$ merges $v_{12}$,
the same could be true of $v_{12}$. Since $v_{12}$ and $v_{22}$ are not
ordered by the ancestor relation, both become LCAs of
$b_3$ and $b$. We observe that this scenario can be prevented if, when
merging $b_1$, $b_3$ insists on an ancestor relation between the last
merged version ($v_{22}$) and the currently merging version.  We call
the last merged version of a branch its \emph{locus}. By maintaining a
total order of among locii of a branch $b$ as new versions are
created, we effectively guarantee the presence of a version (the
locus) that is lower than all the ancestors of the latest version on
$b$. Next, by requiring the locus to be an ancestor of the merging
version, we ensure that all the common ancestors have a single lowest
element, thus enforcing the uniqueness of the LCA.

@@

If $v_1$ and $v_2$ referred in the above definition are the latest
versions on their respective branches $b_1$ and $b_2$, we say $b_2$ is
mergeable into $b_1$, or more generally, $b_1$ and $b_2$ are
mergeable. The definition essentially requires all of $v_1$'s causal
history that precedes its locus $v_o$ to be included in $v_2$'s causal
history, and none of $v_1$'s history that succeeds $v_o$ to be
included in $v_2$'s history. This allows us to view versions $v_1$ and
$v_2$ as having been independently evolved from a common causal
history whose maximum (w.r.t the ancestor relation) is the version
$v_o$. Consequently, $v_o$ becomes the LCA for $v_2$'s merge into
$v_1$. Generalizing this observation for the latest versions on
any two branches, we state the following theorem:

\begin{theorem} [\bfseries Unique LCA]
Every pair of branches in a legal history $H$, if mergeable as per
Def.~\ref{def:mergeability}, have a unique lowest common ancestor.
\end{theorem}

@@

\begin{theorem} [\bfseries Progress]
In a legal branching history $H$ produced by the operational
semantics, if two branches, $b_i$ and $b_j$ are not mergeable (as per
Def.~\ref{def:mergeability}), then there exists a sequence of fork and
merge operations (between mergeable versions) that can be performed on
$H$ to yield a new history $H'$, where $b_i$ and $b_j$ are mergeable.
\end{theorem}
